---
title: "Course 1 - Reproducible Research"
author: "Julio Cesar Melcchioretto"
date: "7/29/2020"

---
```{r}
library(dplyr)
library (ggplot2)
```

## Loading and preprocessing the data
show below the summary of the activity data
```{r}
activity <- read.csv("./activity.csv", sep = ",")
activity$date <- as.Date(activity$date)
summary(activity)
```

## What is mean total number of steps taken per day?
showed below ths histogram of steps per day
```{r}
activity_group <- group_by(activity, date)
total_steps <- summarize(activity_group, steps = sum(steps), na.rm = TRUE)
png("plot1.png")
hist(total_steps$steps, xlab = "total steps per day", main = "Total Steps per day", breaks = 10)
dev.off()
```


the mean of steps per day is:
```{r}
mean(total_steps$steps, na.rm = TRUE)
```

the median of steps per day is:
```{r}
median(total_steps$steps, na.rm = TRUE)
```
## What is the average daily activity pattern?
plot of average number of steps take across the day
```{r}
average_activity <- aggregate(activity$steps, by = list(activity$interval), FUN=mean, na.rm=TRUE)
names(average_activity) <- c("interval", "mean_steps")
png("plot2.png")
plot(average_activity$interval, average_activity$mean_steps, type = "l", xlab = "5-minutes-interval", ylab = "average across all days", main = "average number of steps per interval")
dev.off()
```


the 5 minutes interval which contains the maximum number of steps is:
```{r}
average_activity[which.max(average_activity$mean_steps),]

```
## Imputing missing values
The number of missing data is:
```{r}
sum(is.na(activity))
```

filling all the missing values in the dataset. The NA data is being replaced by mean for the same interval
```{r}
activity_new <- transform(activity, steps = ifelse(is.na(activity$steps), yes = average_activity$mean_steps[match(activity$interval, average_activity$interval)], no = activity$steps))
```

showed below ths histogram of steps per day with the new dataset
```{r}
activity_group_new <- group_by(activity_new, date)
total_steps_new <- summarize(activity_group_new, steps = sum(steps))
png("plot3.png")
hist(total_steps_new$steps, xlab = "total steps per day", main = "Total Steps per day", breaks = 10)
dev.off()
```

the mean of steps per day is:
```{r}
mean(total_steps_new$steps)
```

the median of steps per day is:
```{r}
median(total_steps_new$steps)
```
## Are there differences in activity patterns between weekdays and weekends?

plot of average number of steps take across the day by weekday or weekends
```{r}
activity_new$weekday <- weekdays(activity_new$date)
activity_new <- transform(activity_new, weekday = ifelse(activity_new$weekday =="Saturday" | activity_new$weekday =="Sunday", yes = "Weekend", no = "Weekday"))
final_data <- aggregate(activity_new$steps ~ activity_new$interval + activity_new$weekday, FUN=mean)
names(final_data) <- c("interval", "weekday", "steps")
png("plot4.png")
g<- ggplot(final_data, aes(interval, steps))+geom_line()+facet_grid(.~weekday)
print(g)
dev.off()
```